---
title: "Ravikkumar Rajasingham Capstone 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
Load necessary libraries and packages 
```{r setup, include=FALSE}
install.packages("corrplot", repos = 'http://cran.us.r-project.org')
library(corrplot)
install.packages("ggplot2", repos = 'http://cran.us.r-project.org')
library(ggplot2)
library(gplots)
library(rpart)
install.packages("randomForest", repos = 'http://cran.us.r-project.org')
library(randomForest)
```


Load Dataset into R and add Column Names
```{r}
adult <- read.csv("~/Desktop/adult.data", header=FALSE)
names(adult) <-c("Age", "Workclass", "Fnlwgt", "Education", "Education-Num", "Marital-Status", "Occupation", "Relationship", "Race", "Sex", "Capital-Gain", "Capital-Loss", "Hours-Per-Week", "Native-Country", "Income")
head(adult)

```


Split Data into Training and Testing
```{r}
train_index <- sample(1:nrow(adult), 0.75 * nrow(adult))
training <- adult[train_index,]
testing <-adult[-train_index,]
```

Income Distribution 
```{r}
Income<-table(training$Income)
barplot(Income)
result = summary(training$Income)/nrow(training)*100
result
```


Checking and removing missing values
```{r}
na_count <-sapply(training, function(y) sum(length(which(is.na(y)))))
na_count <-data.frame(na_count)
na_count

na_count_test <-sapply(testing, function(y) sum(length(which(is.na(y)))))
na_count_test <-data.frame(na_count)
na_count_test
```


Re-factorize attributes
```{r}
training$Workclass <- factor(training$Workclass)
training$Education <- factor(training$Education)
training$`Marital-Status` <- factor(training$`Marital-Status`)
training$Occupation <- factor(training$Occupation)
training$Relationship <- factor(training$Relationship)
training$Race <- factor(training$Race)
training$Sex <- factor(training$Sex)
training$`Native-Country` <- factor(training$`Native-Country`)
```

Basic Information of Dataset
```{r}
summary (training)
str(training)
```

Stats for Numerical Attributes
```{r}
hist(training$Age, main = "Age Distribution", xlab = "Age", col = "forestgreen")
boxplot(training$Age, main = "Age")

hist(training$Fnlwgt, main = "Fnlwgt Distribution", xlab = "Fnlwgt", col = "forestgreen")
boxplot(training$Fnlwgt, main = "Fnlwgt")

hist(training$`Education-Num`, main = "Education-Num Distribution", xlab = "Education-Num", col = "forestgreen")
boxplot(training$`Education-Num`, main = "Education-Num")

hist(log10(training$`Capital-Gain`), main = "Capital-Gain", xlab = "Capital-Gain", col = "forestgreen")
boxplot(training$`Capital-Gain`, main = "Capital-Gain")

hist(log10(training$`Capital-Loss`), main = "Capital-Loss", xlab = "Capital-Loss", col = "forestgreen")
boxplot(training$`Capital-Loss`, main = "Capital-Loss")

hist(training$`Hours-Per-Week`, main = "Hours-Per-Week Distribution", xlab = "Hours-Per-Week", col = "forestgreen")
boxplot(training$`Hours-Per-Week`, main = "Hours-Per-Week")
```

Stats for Categorical Attributes
```{r}
tbl <- with(training, table(training$Income, training$Workclass))
barplot(tbl,main = "Workclass Distribution", legend = TRUE, las=2)

tb2 <- with(training, table(training$Income, training$Education))
barplot(tb2,main = "Education Distribution", legend = TRUE, las=2)

tb3 <- with(training, table(training$Income, training$`Marital-Status`))
barplot(tb3,main = "Marital Status Distribution", legend = TRUE, las=2)

tb4 <- with(training, table(training$Income, training$Occupation))
barplot(tb4,main = "Occupation Distribution", legend = TRUE, las=2)

tb5 <- with(training, table(training$Income, training$Relationship))
barplot(tb5,main = "Relationship Distribution", legend = TRUE, las=2)

tb6 <- with(training, table(training$Income, training$Race))
barplot(tb6,main = "Race Distribution", legend = TRUE, las=2)

tb7 <- with(training, table(training$Income, training$Sex))
barplot(tb7,main = "Sex Distribution", legend = TRUE, las=2)

tb8 <- with(training, table(training$Income, training$`Native-Country`))
barplot(tb8,main = "Native Country Distribution", legend = TRUE, las=2)
```

Correlation between Income and Numerical Attributes
```{r}
training$Income <- as.numeric(training$Income)

x <- cor(training[sapply(training, function(x) !is.factor(x))])
corrplot(x, method = "number")
```

Save Training and Testing data
```{r}
write.csv(training, "training.csv", row.names = FALSE)
write.csv(testing, "testing.csv", row.names = FALSE)
```

Log Model
```{r}

model <- glm(Income ~ Age + Workclass + Education + `Marital-Status` + Occupation + Race + Sex + `Hours-Per-Week`, data = training, family = "binomial")
summary(model)

prediction <- predict(model, training, type = 'response')
predict <- rep('<=50K', length(prediction))
predict[prediction>=.5] <- '>50K'
table <- table(predict, training$Income)
table
```

Random Forest Model
```{r}
model2 <- randomForest(Income ~ Age + Workclass + Education + Occupation + Race + Sex, data = training, ntree = 400)
summary(model2)
prediction2 <- predict(model2, training, type = 'class')
prediction2
table2 <- table(prediction2, training$Income)
table2
```

Naive Bayes Model
```{r}
install.packages("naivebayes", repos = 'http://cran.us.r-project.org')
library(naivebayes)
model3 <- naive_bayes(Income ~ Age + Workclass + Education + `Marital-Status` + Occupation + Race + Sex + `Hours-Per-Week`, data = training)
prediction3 <- predict(model3, training, type = 'class')
prediction3
table3 <- table(prediction3, training$Income)
table3
```

